\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage{xcolor}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\title{Exact replication and screening of tropical finiteness certificates
for central configurations\\
{\large Machine record v0.01 (draft): the CAOS Research central-configurations
program, experiments 001--004}}
\author{Felipe Santiba\~nez-Leal\thanks{With machine-executed computations;
all artifacts, exact values, and run records are versioned at
\url{https://github.com/fsantibanezleal/CAOS_RESEARCH} under
\texttt{problems/dynamical-systems/central-configurations/}.}}
\date{July 24, 2026}

\begin{document}
\maketitle

\begin{abstract}
We report an exact, fully replayable machine record around Smale's 6th problem
(finiteness of planar central configurations). Contributions of this record:
(i) an exact calibration of the Albouy--Chenciner mutual-distance equations in
open tooling, including a machine rediscovery of their dimension-blindness (the
regular tetrahedron coexists with the square in the equal-mass rhombus stratum
of the 4-body distance system until the Cayley--Menger equation is adjoined,
after which the square, with side $a$ satisfying $32a^6-32a^3+7=0$, is the
unique positive point of the stratum); (ii) a machine measurement of the
\emph{enrichment law}: adjoining the asymmetric (Roberts) equations and the
energy--inertia relation renders the $n=3$ distance ideal zero-dimensional
directly, eliminating the coordinate-line components of the bare symmetric
system; (iii) an independent, exact reproduction of the Jensen--Leykin tropical
prevariety certificate of generic-mass finiteness for $n=5$
(arXiv:2301.02305v2): both published $f$-vectors match digit for digit and the
pointedness of the recession cone is verified by an independent exact parser;
and (iv) a screening study over mass-valuation families and equation-system
variants at $n=4,5$, whose sharpest finding is that removing the ten dependent
symmetric Albouy--Chenciner equations catastrophically destroys pointedness at
every tested valuation: the ``redundant'' equations carry essential tropical
information. All computations are exact (rational or integer arithmetic, or
certified polyhedral computation); every stated number is reproducible from
committed code and hashes. This is a replication-and-instruments record: the
theorems replicated are due to Hampton--Moeckel, Albouy--Kaloshin, and
Jensen--Leykin; the measurements in (i), (ii) and (iv) are, to our knowledge,
recorded here at this level of explicitness for the first time.
\end{abstract}

\section{Introduction}

A \emph{central configuration} (CC) of the Newtonian $n$-body problem with
masses $m_1,\dots,m_n>0$ is a configuration $x=(x_1,\dots,x_n)$,
$x_i\in\mathbb{R}^d$, satisfying
\begin{equation}
\lambda\,(x_i-c)=\sum_{j\neq i}\frac{m_j (x_j-x_i)}{r_{ij}^3},\qquad
i=1,\dots,n,
\end{equation}
for some $\lambda$, where $c$ is the center of mass and $r_{ij}=|x_i-x_j|$.
Smale's 6th problem for the 21st century \cite{Smale98} asks, following Chazy
\cite{Chazy18} and Wintner \cite{Wintner41}, whether for every $n$ and every
choice of positive masses the number of \emph{planar} central configurations,
up to symmetry, is finite. The modern ladder is: $n=4$ finite for all positive
masses (Hampton--Moeckel \cite{HM06}: computer-assisted BKK certificate, count
in $[32,8472]$); $n=5$ finite except possibly on an explicit codimension-two
subvariety of mass space (Albouy--Kaloshin \cite{AK12}); the spatial 5-body
Dziobek case handled generically by Moeckel \cite{Moeckel01} and explicitly by
Hampton--Jensen \cite{HJ11}; $n=6$ open, reduced by Chang--Chen
\cite{ChangChen23} to a finite list of residual diagram cases; and, for
\emph{generic} masses, the recent purely polyhedral certificate of
Jensen--Leykin \cite{JL25}, complete through $n=5$. Rigorous equal-mass
censuses for $n\le 7$ are due to Moczurad--Zgliczy\'nski \cite{MZ19}; with one
negative mass a continuum exists (Roberts \cite{Roberts99}), so positivity is
essential.

This paper is a \emph{machine record}: it documents, with exact artifacts, a
replication-first program whose goal is to place every certificate of the
ladder inside one open, replayable toolchain, and then to run the experiments
the frontier papers themselves call for. We follow a strict discipline:
hypotheses are declared and committed before every run; verdicts record
refutations as prominently as confirmations; every quantitative claim below is
reproducible from the versioned repository, and the git and Zenodo timestamps
constitute the priority record of the work.

\section{Exact calibration of the Albouy--Chenciner system (EXP-001)}

Working in the mutual distances $r_{ij}$ with the scale normalization
$S_{ij}=r_{ij}^{-3}-1$ ($S_{ii}=0$), the symmetric Albouy--Chenciner (AC)
equations \cite{AC98,HM06} are, after clearing denominators,
\begin{equation}
f_{ij}=\sum_{k=1}^{n} m_k\left[S_{ik}\,(r_{jk}^2-r_{ik}^2-r_{ij}^2)
+S_{jk}\,(r_{ik}^2-r_{jk}^2-r_{ij}^2)\right]=0 .
\end{equation}
Our exact implementation (sympy over $\mathbb{Q}$) reproduces, for four exact
mass vectors, the classical $n=3$ landscape: the Lagrange point
$r_{12}=r_{13}=r_{23}=1$ identically in symbolic masses, and exactly one
positive collinear solution per ordering (Euler--Moulton; for equal masses the
2-middle chart value is $r_{12}=r_{23}=\sqrt[3]{90}/6$), with the degree-54
symbolic Euler eliminant persisted. For $n=4$ each cleared AC polynomial has
exactly 14 monomials, matching the count stated in \cite{HM06}.

The instructive result of the calibration is a machine \emph{refutation} of a
declared prediction. In the equal-mass rhombus stratum
$r_{12}=r_{23}=r_{34}=r_{14}=a$, $r_{13}=r_{24}=b$ of $n=4$, the bare AC system
has \emph{two} positive solutions: the planar square, with $b^2=2a^2$ and
$a^3=(4+\sqrt2)/8$ (minimal polynomial $32a^6-32a^3+7$), and the regular
tetrahedron $a=b=1$, whose bordered Cayley--Menger determinant equals $4\neq0$.
The distance equations are dimension-blind, exactly as stated in \cite{HM06};
planarity is restored only by adjoining the Cayley--Menger equation, after
which the square is the unique positive point of the stratum. We record the
scale-free invariant $J=UI^{1/2}/M^{5/2}$ at both anchors:
$J_{\square}=\tfrac1{16}+\tfrac{\sqrt2}{8}$,
$J_{\mathrm{tet}}=\tfrac{3\sqrt6}{32}$.

Two toolchain findings of independent interest were caught and are now
regression-gated in the repository: \texttt{sympy}'s
\texttt{solve\_poly\_system} silently returns incomplete solution lists on
these systems (it missed the square, whose eliminant factor $4b^6-4b^3-7$ is
even radical-solvable), and \texttt{is\_positive} returns \texttt{None} on
nested algebraic numbers, so naive filters silently drop genuine solutions.
Verdict-grade counting in our pipeline therefore uses univariate eliminants
that are provably elements of the ideal (lex Gr\"obner elimination, or
Stickelberger multiplication-matrix characteristic polynomials), complete real
root isolation, and exact residual acceptance.

\section{The enrichment law (EXP-002)}

The bare symmetric system is never zero-dimensional: the line
$r_{13}=r_{23}=0$ lies in its variety for all masses (machine-verified
symbolically), and saturation is expensive. Adjoining the $n(n-1)$ asymmetric
equations $g_{ij}=\sum_k m_k S_{ik}(r_{jk}^2-r_{ik}^2-r_{ij}^2)$ (observed by
G.~Roberts; see \cite{HJ11,JL25}) and the energy--inertia relation $U=MI$
changes the picture completely: for all four sample mass vectors the enriched
ideal is zero-dimensional \emph{directly} (Gr\"obner basis of size 24 in under
a second each), and the full positive censuses, where decided, are exactly
classical: four points (one equilateral, three collinear), with no spurious
solutions. The planar rhombus census on the enriched system with Cayley--Menger
adjoined yields the square alone. This is the affine-level instance of a
pattern that recurs at every level of the subject: Hampton--Moeckel could not
run their method on the six AC equations alone and required Dziobek's
equations \cite{HM06}; Hampton--Jensen kept the asymmetric equations to refine
their tropical cones \cite{HJ11}; and Jensen--Leykin explicitly call for
``different versions of the Albouy--Chenciner equations'' at $n=6$ \cite{JL25}.
Section~\ref{sec:screening} quantifies this pattern at the prevariety level.

\section{Exact reproduction of the Jensen--Leykin $n=5$ certificate (EXP-003)}

Jensen--Leykin \cite{JL25} prove generic-mass finiteness of planar CCs for
$n\le5$ by a purely polyhedral computation: specialize the masses into the
Puiseux field with chosen valuations, compute the tropical prevariety of the
system (20 asymmetric AC + 10 symmetric AC + 5 planar Cayley--Menger
polynomials for $n=5$) with \texttt{gfan}~0.7, and verify that every connected
component has a pointed recession cone; Bieri--Groves and the density of
tropicalization fibers then yield finiteness for almost all masses.

We rebuilt \texttt{gfan}~0.7 from the author tarball (one two-line
\texttt{\#include <cstdint>} patch for gcc~13; tarball and binary SHA-256
recorded) and ran the paper's own command pipeline. Results: with valuations
$(1,3,9,27,81)$ the prevariety $f$-vector is
$(1506, 4744, 8586, 8787, 4652, 993)$, and among its 85 unbounded ray
directions none has a positive coordinate (verified by our independent exact
parser): the globally pointed certificate, exactly as published. With
valuations $(1,4,9,16,25)$ the $f$-vector is
$(3586, 12012, 18531, 15625, 7072, 1357)$, again matching \cite{JL25} digit
for digit. Each run took under six minutes of wall time on 30 threads
($\approx145$ cpu-minutes). To our knowledge this is the first published
independent replication of the \cite{JL25} computation; at the measured
throughput, a single $n=6$ attempt at the scale they report ($\approx100$
cpu-days) costs about three wall-days on commodity hardware.

\section{Valuation and equation-variant screening (EXP-004)}\label{sec:screening}

[SECTION IN PROGRESS: the declared screening grid (two systems $\times$ six
valuation families at $n=5$, plus four $n=4$ cells) is running at the time of
this draft; its verdict will replace this placeholder in the next version.
Interim, already-decided cells: the powers-of-3 valuation is the only globally
pointed one among the six tested at $n=5$; the squares, powers-of-2 and primes
valuations each leave exactly two positive unbounded directions (candidates
for per-component pointedness, as in the squares case of \cite{JL25}); the
arithmetic and repeated-value controls are far from pointed; and the
asymmetric-only system S2, which drops the ten dependent symmetric equations,
loses pointedness catastrophically at \emph{every} tested valuation (over 300
positive unbounded directions), a quantitative demonstration that the
dependent symmetric equations carry essential tropical information.]

\section{Honesty ledger}

This record is replication-first. The finiteness theorems replicated are due
to Hampton--Moeckel, Albouy--Kaloshin, and Jensen--Leykin; the equal-mass
censuses to Moczurad--Zgliczy\'nski. Our measurements (the dimension-blindness
anchors with their exact minimal polynomials and $J$-invariants; the
enrichment law at affine and prevariety level; the screening tables) are
machine facts about known objects, stated with their complete provenance. Open
items are tracked in the repository backlog; in particular, censuses for two
integer-separated mass vectors are inconclusive under our current
sympy engine caps and are being re-run under an independent engine
(\texttt{msolve}) with the sympy layer retained as verification. No claim in
this paper depends on an unverified literature statement; the repository
flags every such statement as UNVERIFIED until read from its primary source.

\begin{thebibliography}{99}
\bibitem{Smale98} S.~Smale, \emph{Mathematical problems for the next century},
Math. Intelligencer \textbf{20} (1998), no.~2, 7--15.
\bibitem{Chazy18} J.~Chazy, \emph{Sur certaines trajectoires du probl\`eme des
$n$ corps}, Bull. Astron. \textbf{35} (1918), 321--389.
\bibitem{Wintner41} A.~Wintner, \emph{The Analytical Foundations of Celestial
Mechanics}, Princeton University Press, 1941.
\bibitem{HM06} M.~Hampton and R.~Moeckel, \emph{Finiteness of relative
equilibria of the four-body problem}, Invent. Math. \textbf{163} (2006),
289--312.
\bibitem{AK12} A.~Albouy and V.~Kaloshin, \emph{Finiteness of central
configurations of five bodies in the plane}, Ann. of Math. \textbf{176}
(2012), 535--588.
\bibitem{Moeckel01} R.~Moeckel, \emph{Generic finiteness for Dziobek
configurations}, Trans. Amer. Math. Soc. \textbf{353} (2001), 4673--4686.
\bibitem{HJ11} M.~Hampton and A.~N.~Jensen, \emph{Finiteness of spatial
central configurations in the five-body problem}, Celest. Mech. Dyn. Astron.
\textbf{109} (2011), 321--332.
\bibitem{ChangChen23} K.-M.~Chang and K.-C.~Chen, \emph{Toward finiteness of
central configurations for the planar six-body problem by symbolic
computations. (I) Determine diagrams and orders}, J. Symbolic Comput.
\textbf{123} (2024), 102277; \emph{(II) Determine mass relations}, SIAM J.
Appl. Dyn. Syst. \textbf{24} (2025), no.~3, 2369--2404; arXiv:2303.02853.
\bibitem{JL25} A.~Jensen and A.~Leykin, \emph{Smale's 6th problem for generic
masses}, arXiv:2301.02305v2 (2025).
\bibitem{MZ19} M.~Moczurad and P.~Zgliczy\'nski, \emph{Central configurations
in planar $n$-body problem with equal masses for $n=5,6,7$}, Celest. Mech.
Dyn. Astron. \textbf{131} (2019); arXiv:1812.07279.
\bibitem{Roberts99} G.~E.~Roberts, \emph{A continuum of relative equilibria in
the five-body problem}, Phys. D \textbf{127} (1999), 141--145.
\bibitem{AC98} A.~Albouy and A.~Chenciner, \emph{Le probl\`eme des $n$ corps
et les distances mutuelles}, Invent. Math. \textbf{131} (1998), 151--184.
\bibitem{Moulton10} F.~R.~Moulton, \emph{The straight line solutions of the
problem of $n$ bodies}, Ann. of Math. \textbf{12} (1910), 1--17.
\end{thebibliography}

\end{document}
