\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{booktabs}
\usepackage{xcolor}
\definecolor{linknavy}{RGB}{25, 55, 125}
\usepackage[colorlinks=true, linkcolor=linknavy, citecolor=linknavy, urlcolor=linknavy]{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\C}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Gm}{\mathbb{G}_m}
\newcommand{\detJ}{\det JF}

\title{The Jacobian Conjecture after the Three-Dimensional Counterexample:\\
Exact Validation, a General Seed Family, the Geometry of Non-Properness,\\
and Two-Dimensional Equivariant Rigidity}
\author{Felipe Santiba\~nez-Leal\\
\small CAOS Research program, \texttt{github.com/fsantibanezleal/CAOS\_RESEARCH}}
\date{2026-07-21 (working manuscript, v0.05)}

\begin{document}
\maketitle

\begin{abstract}
On 2026-07-19, Levent Alp\"oge announced (crediting the language model Claude Fable 5 and a
question posed by Akhil) an explicit polynomial map $F : \C^3 \to \C^3$ with constant Jacobian
determinant $-2$ and a repeated value, refuting the Jacobian conjecture for every $N \ge 3$. We
report the results of an exact-arithmetic experimental program built on this event. First, we
validate the counterexample independently and compute the full fiber over the announced target:
it consists of exactly the three announced points. Second, we reverse-engineer the structure of
$F$: it is a weighted skew-product over the invariants $v = xy$, $t = x^2 z$, its Keller
property reduces to a two-variable identity, and its fibers are governed by a cubic in a single
composite variable. Third, we derive and verify a general constructor: every admissible triple
(seed polynomial $p$, scale $k$, section $q$) yields a Keller map with determinant $-k\,p(1)^2$
and generic fiber degree $\deg p + 1$; we exhibit new explicit counterexamples with rational
collision certificates in fiber degrees $4$, $5$ and $6$, plus a section-tail instance beyond
the announced family shape, prove the degree law $(5d - 3, 5d - 4, 4)$ within the family, and
show the announced map is its smallest member. Fourth, we make the failure geometry exact: a
preimage escapes to infinity precisely when its fiber root is a multiple root, so the
asymptotic variety of $F$ is the plane $C = 0$ together with the explicit discriminant surface
$27A^2C^2 - 18ABC + 16A + B^3C - B^2 = 0$; over $\mathbb{R}$ this wall separates a three-sheet
from a one-sheet region, and $F$ restricts to a surjective, non-injective real Keller map. The
counterexamples reduce at suitable primes to non-injective Keller maps over $\mathbb{F}_\ell$
of degree below the characteristic. Fifth, we prove that the mechanism cannot reach the
still-open two-variable case: for arbitrary weights, every $\Gm$-equivariant Keller map of
$\C^2$ is linear (a classification with a positive-factor kill at the top degree), so a planar
counterexample, if any, must be genuinely non-equivariant. All computations are exact (rational
or symbolic) and reproducible from the repository; every claim below is labeled as
machine-verified, derived, or conjectural.
\end{abstract}

\section{Introduction}

Let $k$ be a field of characteristic $0$. A polynomial map $F = (F_1, \dots, F_N) : k^N \to k^N$
is a \emph{Keller map} if $\detJ$ is a nonzero constant. The Jacobian conjecture $JC(N)$,
stated in two variables by Kraus in 1884 (with a flawed proof whose gap, control of ramification
at infinity, remained the subject's central obstacle \cite{Rodriguez2026}) and popularized by
Keller in 1939 \cite{Keller1939}, asserted that every Keller map has a polynomial inverse. Over
$\C$, invertibility is equivalent to injectivity, so a refutation requires exactly one repeated
value. The classical positive ladder includes Wang's degree-2 theorem, Moh's verification of
$JC(2)$ up to degree 100, the Bass--Connell--Wright reduction to cubic homogeneous form
\cite{BCW1982}, and Dru\.zkowski's cubic-linear form \cite{Druzkowski1983}; Pinchuk
\cite{Pinchuk1994} refuted the strong real variant by a non-proper local diffeomorphism with
non-constant Jacobian. The conjecture is Problem 16 of Smale's list \cite{Smale1998}.

On 2026-07-19/20, Alp\"oge announced the map, with $u = 1 + xy$,
\begin{equation}\label{eq:F}
F = \bigl(\,u^3 z + y^2 u (4 + 3xy),\;\; y + 3x u^2 z + 3x y^2 (4 + 3xy),\;\; 2x - 3x^2 y - x^3 z\,\bigr),
\end{equation}
with $\detJ \equiv -2$ and
$F(0, 0, -\tfrac14) = F(1, -\tfrac32, \tfrac{13}{2}) = F(-1, \tfrac32, \tfrac{13}{2}) =
(-\tfrac14, 0, 0)$ \cite{Alpoge2026}. Appending identity coordinates refutes $JC(N)$ for all
$N \ge 3$; the case $N = 2$ remains open. At the time of writing no formal paper exists; the
announcement is a machine-checkable certificate, which makes an exact experimental program both
possible and necessary. This manuscript maps the findings of that program (repository:
CAOS\_RESEARCH, problem \texttt{algebraic-geometry/jacobian-conjecture}; experiments cited as
EXP-$nnn$). Each claim is labeled: \textbf{[MV]} machine-verified exact identity,
\textbf{[D]} derived analytically and certified on instances, \textbf{[C]} conjecture.

\section{Exact validation and the full fiber (EXP-001)}

\begin{theorem}[{[MV]}]\label{thm:valid}
For $F$ as in \eqref{eq:F}: $\detJ = -2$ identically; the three announced points are pairwise
distinct with the common image $(-\tfrac14, 0, 0)$; and the fiber
$F^{-1}(-\tfrac14, 0, 0)$ consists of exactly those three points. In particular $F$ is a
non-injective Keller map and $JC(N)$ is false for every $N \ge 3$.
\end{theorem}

The fiber computation is a lex Gr\"obner basis over $\Q$: the eliminant in $z$ is
$\tfrac18(2z - 13)(4z + 1)$, and on the fiber $3x + 2y = 0$, $12y^2 = 4z + 1$. All coefficients
of \eqref{eq:F} are rational, so non-injectivity holds equally over $\mathbb{R}$.

\section{The structure of the counterexample (EXP-002)}

\begin{theorem}[{[MV]}]\label{thm:structure}
$F$ is equivariant for the $\Gm$-action with source weights $(1, -1, -2)$ and target weights
$(-2, -1, 1)$. Writing $v = xy$, $t = x^2 z$, $u = 1 + v$, the reduced data
\[
a_1 = x^2 P, \qquad b_1 = x Q, \qquad c_1 = R / x = 2 - 3v - t
\]
are polynomials in $(v, t)$; each component of $F$ is affine in $z$ with coefficient vector
$(u^3, 3xu^2, -x^3)$; and the generating identity
$a_1 = \tfrac{u b_1}{2} - u^2 + \tfrac{u^3 c_1}{2}$ holds.
\end{theorem}

\begin{lemma}[Skew-product determinant lemma, {[MV]} with generic symbolic coefficients]
\label{lem:skew}
For \emph{any} polynomials $a_1, b_1, c_1$ in $(v, t)$, the map
$(a_1/x^2,\, b_1/x,\, x c_1)|_{v = xy,\, t = x^2 z}$ satisfies
\[
\detJ \cdot c_1^2 = -J_2, \qquad
J_2 = \det \frac{\partial (V', T')}{\partial (v, t)}, \quad V' = b_1 c_1, \; T' = a_1 c_1^2 .
\]
\end{lemma}

For $F$, $J_2 = 2 c_1^2$ \textbf{[MV]}: the Keller property is a \emph{two-variable} identity;
the third dimension only carries weights. Fibers are governed by a single cubic:
with $w = u c_1 / 2$ and $\Phi(w) = w^2 - w^3$,
\begin{equation}\label{eq:fiber}
\Phi(w) = \tfrac14\left(w\, QR - P R^2\right) \quad \text{identically \textbf{[MV]}},
\end{equation}
so preimages of $(A, B, C)$ correspond to roots of $4\Phi(w) - w BC + A C^2 = 0$; at the
announced target this degenerates to $\Phi(w) = 0$ with roots $\{0, 0, 1\}$, matching the three
preimages' $w$-values $(1, 0, 0)$ \textbf{[MV]}. Injectivity dies by \emph{non-properness}: a
Keller map is a local diffeomorphism, so sheets merge only at infinity; this is Pinchuk's
loophole and the modern face of Kraus's gap.

\section{The seed family: a general counterexample constructor (EXP-003/004)}

The identification $c_1 = h(v) - t$ with $h = p/w$ is a degree-2 coincidence: an initial
constructor built on it is \emph{provably unsolvable} for $\deg p \ge 3$ (EXP-003, an exact
unsolvability statement). The failure analysis yields the correct \emph{potential form}.

\begin{theorem}[Constructor v2, {[D]}, certified on instances {[MV]}]\label{thm:family}
Let $p(w) = \sum_{i=1}^{d} p_i w^i$ satisfy
\[
p(0) = 0, \qquad \int_0^1 p = 0, \qquad p(1) \ne 0, \qquad p'(1) \ne 2 p(1),
\]
let $k \ne 0$, $m = -k\,p(1)$, and let $q(v)$ be any polynomial with $q(0) = k$ and
$q'(0) = k\,(p(1) - p'(1))/(p'(1) - 2p(1))$, with arbitrary higher-order tail. Set
$\Phi(w) = \int_0^w p$, $s = c_1 = q(v) - t$, $w = u s / k$, and define
\[
V' = k^2 p(w) + m s, \qquad T' = w V' - k^2 \Phi(w), \qquad
b_1 = V'/s, \qquad a_1 = T'/s^2 .
\]
Then $a_1, b_1$ are polynomials in $(v, t)$, and
$F_{p, k, q} = (a_1 / x^2,\; b_1 / x,\; x c_1)$ is a polynomial Keller map of $\C^3$ with
\[
\det J F_{p,k,q} = -\frac{m^2}{k} = -k\,p(1)^2 \quad (\text{constant, for every admissible } (p,k,q)),
\]
generic fiber degree $d + 1$, and the exact reconstruction inverse: from a root $w$ of the fiber
polynomial $k^2\Phi(w) - (w\,BC - AC^2)$,
\[
s = \frac{BC - k^2 p(w)}{m}, \quad u = \frac{k w}{s}, \quad x = \frac{C}{s}, \quad v = u - 1,
\quad t = q(v) - s, \quad y = \frac{v}{x}, \quad z = \frac{t}{x^2}.
\]
The announced map \eqref{eq:F} is the instance $p = 2w - 3w^2$, $k = 2$, $q$ affine; at $d = 2$
this shape is rigid up to the scale $k$ \textbf{[MV]}.
\end{theorem}

The three admissibility conditions are exactly the low-weight polynomiality constraints at the
origin; the potential form makes $J_2 = m^2 s^2 / k$ automatic, so \emph{the Keller property
costs nothing}: this is the structural reason the conjecture is false in dimension $\ge 3$.

\begin{theorem}[New explicit counterexamples, {[MV]}]\label{thm:instances}
The following maps are non-injective polynomial Keller maps of $\C^3$, each certified by an
engineered rational collision:
\begin{center}
\begin{tabular}{lllll}
\toprule
Seed $p$ & $k$ & $\det$ & Component degrees & Generic fiber degree \\
\midrule
$w - 2w^3$ & $3$ & $-3$ & $(12, 11, 4)$ & $4$ \\
$8w - 12w^2 + 4w^3 - 5w^4$ & $14$ & $-350$ & $(17, 16, 4)$ & $5$ \\
$2w - 3w^2$, section tail $v^2$ & $2$ & $-2$ & $(8, 7, 5)$ & $3$ \\
\bottomrule
\end{tabular}
\end{center}
For the first: $(-\tfrac{1}{15}, 18, 5130)$ and $(\tfrac{1}{24}, -18, -10368)$ both map exactly
to $(-54, -54, 1)$. Collisions are engineered by prescribing two rational roots $w_1 \ne w_2$ and
solving the linear system $w_i \tilde V - \tilde T = k^2 \Phi(w_i)$ for the target invariants.
\end{theorem}

The section-tail instance changes the component-degree pattern at fixed fiber degree and
determinant: the family is strictly larger than the seed-only family circulating informally.
Every negative rational number is realized as a Keller determinant within the family.

\begin{theorem}[Degree laws and family-internal minimality, {[MV]} for $d \le 5$]\label{thm:degrees}
For affine sections $q$ the component total degrees of $F_{p,k,q}$ are exactly
$(5d - 3,\, 5d - 4,\, 4)$, $d = \deg p$ (verified $d = 2, \dots, 5$; the $d = 5$ instance, seed
$w - 3w^5$, $k = 5$, $\det = -20$, degrees $(22, 21, 4)$, fiber degree 6, carries its own
rational collision certificate). The scale $k$ never changes degrees; section tails strictly
increase them; degree-1 seeds are impossible (the integral condition forces $p = 0$), so the
fiber-degree floor is 3. Consequently the minimal degree vector over the family is $(7, 6, 4)$,
attained exactly at the announced map: \emph{the July 2026 counterexample is the smallest
instance of its own family}. Globally, only Wang's degree-2 theorem constrains $\C^3$; the
existence of a counterexample of total degree $3$ to $6$ remains open.
\end{theorem}

\begin{remark}[Characteristic $p$, {[MV]}]
The $d = 3$ instance reduces mod $\ell = 13$ and $\ell = 101$ to explicit non-injective Keller
maps over $\mathbb{F}_\ell$ of total degree $12 < \ell$ (the rational collision reduces to
distinct points with a common image, e.g. mod 13: $(6,5,8)$ and $(6,8,6)$ both to $(11,11,1)$).
Unlike the classical char-$p$ example $x - x^p$ (degree $= p$), the degree sits strictly below
the characteristic; expected from standard integer-model reductions, now explicitly certified.
\end{remark}

\section{The asymptotic variety: where sheets merge}

\begin{theorem}[Escape $=$ multiple fiber root, {[MV]} on $F$ and the $d=3$ instance]
\label{thm:asymptotic}
In the potential form, the fiber polynomial $W(w) = k^2\Phi(w) - wBC + AC^2$ satisfies
$W'(w) = k^2 p(w) - BC$ identically, and the reconstruction scale is $s = -W'(w)/m$. Hence a
fiber root fails to reconstruct a finite preimage (the preimage escapes to infinity) precisely
when it is a \emph{multiple} root of $W$: over $C \ne 0$ the non-properness locus is the
discriminant locus of $W$. For the announced $F$,
\[
A(F) \;=\; \{C = 0\} \;\cup\; \{\, 27 A^2 C^2 - 18 A B C + 16 A + B^3 C - B^2 = 0 \,\}.
\]
Verified end to end: the target $(0, 1, 1)$ lies on the surface and its exact fiber is the
single point $(2, -\tfrac12, \tfrac98)$ (two sheets escaped); over generic $(A, B, 0)$ the
fiber is the single flat-sheet point $(0, B, A - 4B^2)$; the collision target $(-16, -16, 1)$
lies off the surface with a full 3-point rational fiber. Collisions are generic; escapes are
codimension 1. Kraus's ``ramification at infinity'' is, for this family, the discriminant of a
univariate polynomial.
\end{theorem}

\section{The two-dimensional frontier: obstruction, rigidity, real census (EXP-005/006/010/011)}

\begin{theorem}[{[MV]} with generic symbolic $f, g$]\label{thm:2d}
Let $\Gm$ act on $\C^2$ with weights $(1, -a)$, $a \ge 1$, and let $F = (x^{b_1} f(v),\,
x^{b_2} g(v))$, $v = x^a y$, be an equivariant polynomial map. Then
\[
\detJ = x^{\,b_1 + b_2 + a - 1}\left(b_1 f g' - b_2 f' g\right),
\]
so the Keller condition forces $b_1 + b_2 = 1 - a$ and the single univariate identity
$b_1 f g' - b_2 f' g = \mathrm{const} \ne 0$.
\end{theorem}

\begin{proposition}[One-invariant collapse]\label{prop:collapse}
If the reduced data $(V', T')$ of a weighted skew-product depend on a single variable, then
$dV' \wedge dT' = 0$, hence $J_2 \equiv 0$ and $\detJ \equiv 0$: no Keller map arises. Since a
nontrivial $\Gm$-action on $\C^2$ has one independent weight-0 invariant (against two in
$\C^3$), the potential-form mechanism of Theorem~\ref{thm:family} is structurally unavailable in
dimension $2$.
\end{proposition}

Additionally \textbf{[MV]}: no coordinate slice of \eqref{eq:F} (each variable frozen at rational
test values, any output pair) is a 2D Keller map.

\begin{theorem}[2D equivariant rigidity, {[D]}, every mechanical step {[MV]} on sweeps]
\label{thm:rigidity}
For any $\Gm$-action on $\C^2$ with weights $(w_1, -w_2)$, $w_1 \ge 1$, $w_2 \ge 0$, every
equivariant polynomial Keller map is linear: $(f_0 x,\, g_0 y)$ or $(f_0 y,\, g_0 x)$.
\end{theorem}

\begin{proof}[Proof shape]
Equivariant components are $P = x^{i_1} y^{j_1} f(v)$, $Q = x^{i_2} y^{j_2} g(v)$ with
$v = x^{w_2} y^{w_1}$, and the determinant factors as
\[
\detJ = x^{i_1 + i_2 - 1}\, y^{j_1 + j_2 - 1}
\bigl[ (i_1 j_2 - i_2 j_1)\, f g + v\,(\beta_1 f g' - \beta_2 f' g) \bigr],
\qquad \beta_r = w_1 i_r - w_2 j_r
\]
(certified on 648 weight-times-exponent cases with generic symbolic $f, g$). Constancy forces
$i_1 + i_2 = j_1 + j_2 = 1$, and the mixed shapes ($P = xy f$, $Q = g$ and its swap) admit no
Keller solution (their bracket is divisible by $v$; empty solves). The surviving shapes reduce
to $fg + v(w_1 f g' + w_2 f' g) = c$ (or its swap), whose coefficient at the top bidegree
$N = d_f + d_g$ is $(1 + w_1 d_g + w_2 d_f)\, f_{d_f}\, g_{d_g}$ with a strictly positive
factor: the leading coefficients die and by descent $f, g$ are constant (certified: empty
constrained solves for all exact bidegrees $\le 5$ over eight weight pairs).
\end{proof}

The independent enumeration EXP-006 (216 polynomial Keller instances over a bounded window, all
648 exactly-computed in-image fibers single-point, every instance linear \textbf{[MV]}) is the
experimental companion of Theorem~\ref{thm:rigidity}. Consequence: a two-variable
counterexample, if one exists, must be genuinely non-equivariant: the symmetry class behind the
entire three-dimensional family is completely closed in dimension 2.

\begin{theorem}[Real fiber census of $F$, {[MV]}]\label{thm:realcensus}
Over $\mathbb{R}$, the announced $F$ restricts to a surjective, orientation-reversing
($\det = -2$), non-proper, non-injective real Keller map. For real targets off the
discriminant wall the number of real preimages is 3 (region $D > 0$, which contains the
collision targets) or 1 (region $D < 0$); two independent exact routes (Sturm isolation of the
fiber cubic plus reconstruction, versus Gr\"obner elimination) agree at all samples, and no
empty real fiber exists on a 36-target rational grid (an odd-degree argument makes this
structural). In particular the constant-Jacobian \emph{real} statement in dimension 3 falls
with the same example; Pinchuk's planar counterexample (1994) concerned the non-constant case.
\end{theorem} The honest map of remaining
routes to a hypothetical 2D counterexample: no symmetry at all; non-reductive
($\mathbb{G}_a$) structures, which face the same one-invariant collapse in naive form; or
gradings at infinity in the Newton-polygon/Abhyankar sense. Conversely, a proof of $JC(2)$ in
this program's terms is a properness statement: the reconstruction inverse of
Theorem~\ref{thm:family} shows exactly how asymptotic values are produced in dimension 3
(denominators $BC = k^2 p(w)$), and whether one-invariant geometry leaves room for such escape
is, in our view, the sharpest transferable open question.

\section{Discussion and program}

\textbf{Why the conjecture is false.} In the equivariant class, being Keller is a two-variable
potential identity, satisfiable for free at any fiber degree $\ge 3$; nothing in the
determinant sees properness. \textbf{What is essential.} Two independent weight-0 invariants;
the potential form; valuation conditions at the origin. \textbf{Consequence cascade
(verified from primary sources, EXP-016).} The Mathieu conjecture is false for $SU(3)$ (Mathieu
1997: Mathieu for $SU(N)$ implies $JC(N)$; abelian case Duistermaat--van der Kallen 1998); the
Gaussian moments conjecture is false (GMC implies JC, Derksen--van den Essen--Zhao 2017); Zhao's
vanishing conjecture is false (equivalent to JC); the Image conjecture is false in some
dimension (implies the vanishing conjecture); and the full Dixmier conjecture is false (stable
equivalence $JC(2n) \Leftrightarrow \mathrm{Dixmier}(n)$, Belov-Kanel--Kontsevich and
Tsuchimoto: $JC(4)$ false gives $\mathrm{Dixmier}(2)$ false; $\mathrm{Dixmier}(1)$ remains
open). \textbf{Next experiments.} EXP-010 (2D widening:
general weight pairs, higher degrees); the global-minimality search in degrees 3 to 6
(GPU-widened ansatz enumeration); the Bass--Connell--Wright push-through of $F$ to explicit
cubic form; primary-source verification of the consequence cascade.


\subsection*{The weighted landscape: uniqueness of the mechanism (EXP-012)}
Generalizing to source weights $(1, -1, -m)$, the determinant lemma
$\detJ = x^{\sigma+m} D$ and the pairing reductions $J_2 = \pm c_1^{\,r_a+r_b-1} D$ hold
generically \textbf{[MV]}, and the reduced Jacobian vanishes to order $m-1$ along the section.
The survey outcome: at $m = 1$ the reduced pair is itself a planar Keller map, so that class is
a \emph{bridge} to $JC(2)$ (93 scanned instances, all injective; a collision there would refute
$JC(2)$); at $m = 3$ the class $o = (-2,-2,1)$ is \emph{empty} (a one-line valuation proof,
machine-certified) and $o = (-3,-1,1)$ is rigid in scans (60 instances); at $m = 4$ the
potential-form family does \emph{not} exist: with seed degree up to 5 the admissibility system
has Gr\"obner basis $[1]$ under $\mu \ne 0$ \textbf{[MV]}. Our design hypothesis (a parity law
producing families at every even $m$) was thereby refuted by its own experiment; the finding is
stronger: \emph{within the scanned weighted landscape, the announced counterexample's weight
system is the unique collision-capable mechanism}. Scope: sparse bounded scans, quantified in
the experiment record; wider weight systems and higher-degree seeds are queued.

\subsection*{Reproducibility}
Every theorem labeled [MV] is an exact sympy computation over $\Q$ with a persisted script and
output in the repository (experiments EXP-001 through EXP-012 under
\texttt{problems/algebraic-geometry/jacobian-conjecture/}); each experiment declares its
hypothesis before the run and records an adversarial-validation route with its verdict.

\begin{thebibliography}{9}
\bibitem{Alpoge2026} L. Alp\"oge, announcement of the counterexample, X (@\_\_alpoge\_\_),
2026-07-20; question posed by Akhil, map produced with Claude Fable 5. Independent press
coverage: M. Sparkes, \emph{New Scientist}, 2026-07-20.
\bibitem{BCW1982} H. Bass, E. H. Connell, D. Wright, The Jacobian conjecture: reduction of
degree and formal expansion of the inverse, \emph{Bull. Amer. Math. Soc. (N.S.)} \textbf{7}
(1982), 287--330.
\bibitem{Druzkowski1983} L. M. Dru\.zkowski, An effective approach to Keller's Jacobian
conjecture, \emph{Math. Ann.} \textbf{264} (1983), 303--313.
\bibitem{Keller1939} O.-H. Keller, Ganze Cremona-Transformationen, \emph{Monatsh. Math. Phys.}
\textbf{47} (1939), 299--306. doi:10.1007/BF01695502.
\bibitem{Pinchuk1994} S. Pinchuk, A counterexample to the strong real Jacobian conjecture,
\emph{Math. Z.} \textbf{217} (1994), 1--4.
\bibitem{Rodriguez2026} L. O. Rodr\'iguez D\'iaz, On the origin of the Jacobian conjecture,
\emph{C. R. Math.} \textbf{364} (2026), 363--370. arXiv:2512.23614.
\bibitem{Smale1998} S. Smale, Mathematical problems for the next century, \emph{Math.
Intelligencer} \textbf{20} (1998), 7--15.
\bibitem{vdEssen2000} A. van den Essen, \emph{Polynomial Automorphisms and the Jacobian
Conjecture}, Progress in Mathematics 190, Birkh\"auser, 2000.
\end{thebibliography}

\end{document}
